---
title: "`r params$title`"
subtitle: "`r paste0('Generated with mutscan v', packageVersion('mutscan'))`"
date: "`r Sys.time()`"
output: 
    html_document:
        theme: united
        toc: true
        toc_float: true
        code_folding: hide
params: 
    se: "unspecified"
    title: "mutscan QC report"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
    library(SummarizedExperiment)
    library(mutscan)
    library(DT)
    library(dplyr)
})
```

# Data object

The input data use in this report is contained in the following 
SummarizedExperiment object:

```{r, class.source = "fold-show"}
params$se
```

# Sample annotation and filtering summary

The table below summarizes the available sample information, including 
annotations provided by the data analyst as well as filtering information 
collected by `mutscan`. 

```{r, warning = FALSE}
tmpdf <- as.data.frame(colData(params$se))
extradf <- do.call(
    dplyr::bind_rows, 
    lapply(names(metadata(params$se)$parameters), function(nm) {
        data.frame(Name = nm, 
                   lapply(metadata(params$se)$parameters[[nm]], 
                          function(x) paste(x, collapse = "; ")))
    })
)
DT::datatable(dplyr::full_join(tmpdf, extradf, by = "Name"),
              extensions = "FixedColumns",
              rownames = FALSE,
              options = list(
                  paging = TRUE, searching = TRUE, info = FALSE,
                  pageLength = 20, 
                  sort = TRUE, scrollX = TRUE, fixedColumns = list(leftColumns = 1)
              )) %>%
    formatStyle("Name", "vertical-align" = "center")
```

The plots below focus on the filtering information, and show the number of 
reads remaining after each filtering step performed by `mutscan`, as well as the 
fraction of reads that are filtered out by each step.

```{r, fig.height = 1.5 * ncol(params$se), fig.width = 10}
tryCatch({mutscan::plotFiltering(params$se, valueType = "reads", onlyActiveFilters = TRUE, 
                                 displayNumbers = TRUE, plotType = "remaining", facetBy = "sample")},
         error = function(e) message("Couldn't generate filtering plots."))
```

```{r, fig.height = 1.5 * ncol(params$se), fig.width = 10}
tryCatch({mutscan::plotFiltering(params$se, valueType = "fractions", onlyActiveFilters = TRUE, 
                                 displayNumbers = TRUE, plotType = "filtered", facetBy = "step")},
         error = function(e) message("Couldn't generate filtering plots."))
```

# Pairs plot

Next, we show a pairs plot, displaying the correlation between each pair of
samples. 

```{r, fig.height = 10, fig.width = 10}
tryCatch({mutscan::plotPairs(params$se)},
         error = function(e) message("Couldn't generate pairs plot."))
```

# Total count plot

The plot below shows the total count for each sample, across all features
in the SummarizedExperiment object.

```{r}
tryCatch({mutscan::plotTotals(params$se, selAssay = "counts")},
         error = function(e) message("Couldn't generate total count plot."))
```

# Distribution plots

The next group of plots displays the distribution of counts for each sample 
(across all features in the SummarizedExperiment object). The same information 
is displayed as a 'knee plot', where the counts for each sample are 
arranged in decreasing order, and as a density plot. Note that both plots 
display the values on a log scale; hence any zero values are not shown. 

```{r, warning = FALSE}
tryCatch({mutscan::plotDistributions(params$se, selAssay = "counts", pseudocount = 1,
                                     groupBy = NULL, plotType = "knee", facet = FALSE)},
         error = function(e) message("Couldn't generate distribution plot."))
```

```{r, warning = FALSE}
tryCatch({mutscan::plotDistributions(params$se, selAssay = "counts", pseudocount = 1, 
                                     groupBy = NULL, plotType = "density", facet = FALSE)},
         error = function(e) message("Couldn't generate distribution plot."))
```

# Session info

<details>
<summary><b>
Click here to see the package versions used to generate this report.
</b></summary>
```{r}
sessionInfo()
```
</details>

